Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2015-11-11 reviews/response-marshall.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.341,0.596,0.824}
\definecolor{darkred}{rgb}{0.439,0.157,0.145}  % from Advances in Geoscience

\begin{document}
\textbf{Authors' response to S.~J.~Marshall}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{darkred}{TODO: #1}}

To S.~J.~Marshall,

% ----------------------------------------------------------------------

\sechead{1 \quad Summary comments}

\referee{%
    Seguinot and colleagues provide the first detailed glaciological modelling
    of that I am aware of for the Cordilleran Ice Sheet in western North
    America, making this a novel and long overdue contribution. The authors
    have not only made new advances with this contribution, they have done so
    in an impressive leap forward. This is an excellent and carefully-presented
    study which is likely to rejuvenate interest and debate in Cordilleran Ice
    Sheet reconstructions. The balance between numerical modelling and glacial
    geological/geomorphological considerations is unusually strong, and the
    authors can be commended for this emphasis. This adds tremendous value to
    the results and increases confidence in the modelling, and I also
    appreciate that the authors point out areas where the numerical model is
    not in accord with the geological record.}

\referee{%
    The manuscript is well-written and beautifully illustrated, and I have very
    few substantive comments. The choices made by the authors are logical and
    well-explained, and they reach several well-substantiated conclusions: a
    two-phase Cordilleran glaciation, a reasonably robust estimate of CIS
    volume at LGM, the general model of CIS growth through multiple alpine
    icefields, and the importance of the Skeena Mountain inception centre. One
    can always quibble with specific aspects of the model design and climate
    scenarios, but the authors have explored a reasonable span of `solution
    space' and these aspects of the Cordilleran ice sheet history appear to be
    robust features of the simulations.}

\referee{%
    The modelling strategy and results presented here stand to be widely cited,
    and I expect that it will serve as a springboard for additional studies
    from others in the international community. I recommend this manuscript for
    publication in The Cryosphere without reservations.}

% ----------------------------------------------------------------------

\sechead{1 \quad Specific comments}

\referee{%
    The Cordilleran Ice Sheet is difficult to model due to its complex
    topography and multiple inception centres (and possibly multiple
    domes/divides), strong regional climatic gradients, which require
    relatively high-resolution climate input fields, and a dearth of
    paleoclimate proxies for western North America to inform spatial and
    temporal variations in climate conditions during the glacial period. The
    authors confront these challenges well, with a adequate ice sheet model
    resolution (5 to 10\,km) and ice physics, carefully calibrated `control'
    climatology (published in Seguinot et al., 2014), and a good exploration of
    different paleoclimate time series histories in this contribution.}

\referee{%
    Nevertheless, it is not clear that ice-core based paleoclimate proxies from
    Greenland or Antarctica are appropriate for western North America. This may
    be particularly true of Greenland proxies, where the amplitude of D-O
    (millennial) climate variations is exceptionally strong and is likely to be
    regional. Because these remote ice core records are `scaled' based on only
    one constraint, producing a CIS maximum configuration that resembles the
    geological record, it is difficult to assess the pre-LGM simulations or the
    details of the modelled ice divide structure, ice thickness, etc. The
    robustness of the conclusion that Greenland and Antarctic ice core records
    are good proxies of glacial climate variability in western North America is
    therefore not so clear, but it is admittedly hard to do better at this
    time. I do wonder if there is any hope from more regional climate proxies
    such as the Logan ice cores or the off-shore Vancouver Island sediment
    records that are cited from Cosma et al. This is worth a short discussion.}

\todo{%
    Logan ice core covers only the last 16\,Ka and is not a good proxy for
    temperature. The Vancouver SST record covers only deglaciation. Check new
    papers for updates. Discuss in the manuscript.}

\referee{%
    Similarly it is difficult to know the errors and uncertainties associated
    with the assumption of fixed modern-day spatial patterns for temperature
    and precipitation. I suspect that the sensitivity of this assumption far
    exceeds that associated with the different paleoclimate proxies. Such that,
    for example, one could readily imagine different assumptions, such as a
    maritime effect that gives reduced glacial cooling near the coast vs. in
    the interior, that is a stronger effect than the difference between
    different paleo-climate proxies with respect to the timing of LGM, ice
    divide structure, etc. But this is a very reasonable start, what the
    authors have done -- there is always going to be more parameter space to
    explore in future studies. As above, I would perhaps just suggest a small
    discussion of the authors' opinion on this question, the uncertainty or
    possible influence of this assumption of modern-day climate patterns.}

\todo{Discuss in the manuscript.}

\referee{%
    Several minor points and grammatical corrections are included in the
    attached text. Nothing that will require much thought -- this is a really
    impressive piece of research, overall, and I am hard-pressed to find any
    criticism of it. It is one of the easiest reviews I have ever done.
    Congratulations to the authors and thanks for this fine work.}

\todo{Correct mistakes.}

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------

\end{document}

### 2015-11-11 reviews/response-stumpf.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}

\begin{document}
\textbf{Authors' response to A.~Stumpf}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

To A.~Stumpf,

\referee{%
    Seguinot et al. present a well designed numerical model for the Cordilleran
    Ice Sheet (CIS) in North America for the last two glaciations occurring
    over the past 120000 years. Although the constraints on such a model are
    not yet all fully understood, their proposed simulations attempt to take
    account all the complexity of the glacial system, and utilizes a variety of
    data sets. The modeling confirms what geologists have observed in the
    field; there was a fast decay of the CIS during MIS~2 and MIS~4 and
    non-glacial conditions existed during MIS (Olympia Nonglacial Interval;
    e.g., Plouffe and Jett\'e, 1997). I applaud them for undertaking of such a
    difficult task, and notifying the reader where input data is sparse or
    inconclusive.}

\referee{%
    I provide the following general comments and observations which may help
    the authors in revising the manuscript for final publication. Many of these
    points are both my personal suggestions and also the recommendations of
    others currently researching the CIS or who have completed studies in the
    past.}

\referee{%
    1) With the large amount of research that has been undertaken to map the
    landforms and deposits of the CIS by federal and provincial scientists,
    academic faculty, and undergraduate and graduate students and determine the
    extent, volume and dynamics ice sheet, the impression left on the reader by
    the opening sentence would be incorrect. These studies have greatly
    advanced our understanding of the CIS, and could be an important dataset to
    test against the modeling. Stumpf et al. (2014) provides a list of some of
    these studies.}

\referee{%
    2) To help the reader better understand the maps presented, I would
    recommend some spatial information be added (e.g., latitude/longitude
    grids; political boundaries; lakes and rivers, place names etc..).}

\referee{%
    3) Was the model tested against regional-scale ground-based data (e.g.,
    Ferbey et al. 2013) to constrain ice divide positions, ice flow direction,
    and ice sheet thickness?}

\referee{%
    4) For the central sector of the CIS, Stumpf et al. (2000) provides some
    insight into the chronology and effectiveness of glacial erosional during
    the MIS. In figure~5, and in the accompanying text, they describe how
    landforms on the surface formed. For example, in lake valleys east of the
    Skeena Mountains, it appeared the major glacial streamlined landforms were
    formed during a longer glacial advance phase, with ice flow paralleling the
    valleys, and later flows, some perpendicular the valley flow, only weakly
    impacted them.}

\referee{%
    5) Stumpf et al. (2000) was the first study to extensively document a
    predominant westerly directed ice flow across high elevations in the Skeena
    and Coast Mountains. This flow direction appeared to continue into the
    late-glacial period. Other subsequent studies also confirm that
    late-glacial readvance eastward out of these mountains and retreat of ice
    margins westward into these mountains was limited.}

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------

\end{document}

### 2015-11-12 reviews/response-jarosch.tex

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}

\begin{document}
\textbf{Authors' response to A.~H.~Jarosch}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\doi}[1]{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}
\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

@@

\sechead{1 \quad General comments}

\referee{%
    Seguinot et al. present in this well written and structured manuscript a
    numerical modelling study of the Cordilleran ice sheet through the last
    glacial cycle. The model is driven by several temperature reconstructions
    based on proxy data and model output is subsequently compared in detail to
    the existing geological evidence in the region. The study is of significant
    relevance as it focuses on the Cordilleran ice sheet evolution in the past,
    which is still poorly understood.}

\referee{%
    Nevertheless, the manuscript is quite unbalanced in its presentation as it
    focuses strongly on section~4 (Comparison with geological record) and by
    doing so neglects crucial details in section~2 (Model setup). This poses a
    fundamental challenge for understanding the science presented. If it is not
    quite clear what the model does and how it performs to start with, it
    becomes difficult to discuss the results of the modelling study and why
    there are mismatches with geological evidence.}

\referee{%
    An overall sensitivity study of the parameters used in the model is
    completely lacking, thereby making it almost impossible to understand
    different responses of the ice sheet model. After reading the manuscript,
    one is left with the impression that the authors assume the PISM ice sheet
    model to be a black box which just requires one initial ``correct'' setup
    with literature values. This notion is reflected in the current manuscript,
    where almost all mismatches of model output with geological evidence (as
    discussed in section~4) are attributed to climate variations lacking in the
    proxy data, or climate-ice sheet feedback mechanisms not represented in the
    model chain. Similarly in a previous study \citet{Seguinot.etal.2014} have
    focused only on the driving climate sensitivities and have omitted
    influences of the ice sheet model as well as mass balance model pa-
    rameters even though they note in that study that these sensitivities
    require attention as well.}

@@

% ----------------------------------------------------------------------

\sechead{2 \quad Specific Comments}

\referee{%
    I refer to text locations in the discussion paper by page number (P) and
    line number range (L) for the specific comments.}

\referee{%
    \textbf{P4151 L11-16:}
    In this sentence the authors refer back to their previous work
    \citep{Seguinot.etal.2014} and highlight that the NARR temperature and precipitation
    fields are the most suitable present day climate datasets to be used.
    Especially since the NARR precipitation fields include steep precipitation
    gradients which are required as identified by et \citet{Seguinot.etal.2014}.
    NARR is delivered on a 32\,km Lambert grid, and thus it is ques- tionable
    how ``steep'' these gradients can be, given the rather smooth
    representation of the existing topography on a 32\,km grid. \citet{Seguinot.etal.2014}
    have partly discussed that however. NARR precipitation and
    temperature fields have been evaluated in detail based on available station
    data for large parts of the study domain dealt with in this manuscript.
    This evaluation \citep{Jarosch.etal.2010} demonstrated that NARR has
    difficulties simulating orographic processes in the Coast Mountains which
    in turn results in unrealistic atmospheric conditions over the Rocky
    Mountains. \citet{Jarosch.etal.2010} further concluded that physics based
    downscaling is required to adequately drive glacier models in that region.
    The authors should argue in more detail here why they think that NARR
    precipitation fields at 32\,km are adequate to drive their model and
    reflect their arguments with the findings of \citet{Jarosch.etal.2010}. A
    solid argument here is of special importance as the authors assume the
    present day precipitation fields to be valid throughout their model time
    period (120\,ky to present) without further corrections (cf. section~2.4
    equation~6).}

\referee{%
    \textbf{P4152 L11:}
    Basal topography is ``derived'' from ETOPO1 data. What does this mean? Do
    the authors just re-sample the DEM data to their 10\,km and 5\,km model
    grids (P4152 L21-22) or is there more processing done? The ETOPO1 data
    contains the present day ice volumes within the study region.
    \citet{Clarke.etal.2013}
    have estimated the ice volume in parts of that region to be $2530
    \pm 220\,km^2$, with maximum ice thicknesses up to 200\,m. It can be argued
    that the volume is negligible in this study and the authors should do so if
    they think this is appropriate, but I wonder about the ice thicknesses.
    Assuming that the authors did not remove the present day ice cover, basal
    topography could be up to 200\,m higher that it actually is in reality.
    Given their used temperature lapse rate of $6\,K\,km^{-1}$ (P4157 L1),
    parts of the Cordilleran ice sheet growing in those regions with 200\,m too
    high topography would experience a 1.2\,K colder atmosphere than it
    actually should in reality. This favours unrealistic ice growth and thus
    the omission of present day ice cover removal should be clearly argued for
    in the manuscript.}

@@

\referee{%
    \textbf{P4153 L2-3:}
    That the ``shallow shelf approximation'' (SSA) is used as a ``sliding law''
    for the ``shallow ice approximation'' (SIA) is a confusing statement in
    this context. Bueler has coined the term in his 2009 paper as cited in the
    manuscript. However the casual reader will be confused at this point,
    especially since the authors state the pseudoplastic sliding law the model
    actually uses in equation~1. I would recommend to leave out the statement
    on the SSA being the ``sliding law'' for the SIA.}

\referee{%
    \textbf{P4153 L5-6:}
    As stated here, ice rheology within the used ice sheet model is based on
    \citet{Aschwanden.etal.2012}. This enthalpy based formulation has proofed
    itself to be very suitable for estimating ice rheology in ice sheet models,
    but it also depends on several parameters to translate enthalpy within the
    ice to ice viscosity \citep[see][equations 62-65)]{Aschwanden.etal.2012}. The
    authors do not mention any of these parameters (e.g. any of the rate
    factors or nonlinear power $n$ ) within the manuscript or in Table~1. I
    have mentioned above in the general comments section that parameters used
    in ice rheology and basal sliding formulations are important model
    parameters which will influence the ice sheet model output and that a basic
    sensitivity study on those parameters is required to understand the model
    results. Here the authors could start with listing the parameters used in
    the ice rheology formulation, than continue with estimating uncertainties
    for those from literature and afterwards perform additional model
    simulations to identify the influence of the chosen parameter sets on the
    ice volume and ice margin position history the model creates. In the end
    the authors will be able to identify the relative importance of
    uncertainties in driving climate as well as model parameters, which will
    strengthen their discussion in section~4.}

@@

\referee{%
    \textbf{P4153 L8:}
    It is not clear where the geothermal heat flux boundary is located. Does
    the ``depth of 3\,km'' refer to a depth measured from the ice surface,
    which would not make much sense for a ice thickness evolving ice sheet
    model or is it measured from the ice-bedrock interface downward. In that
    case the term ``computed subglacially'' is confusing as it refers to the
    ice-bedrock interface. Please be more specific here.}

\referee{%
    \textbf{P4153 L16 - P4154 L0:}
    Here the authors describe the basal sliding setup in their model. However
    they do not explain how they came up with the parameters used in
    equations~1-3 that are listed in Table~1 (part on ``basal sliding''). What
    motivates these parameter choices (references?) and how sensitive is the
    model and its results to these choices? Both question come instantly to
    mind and need to be addressed in detail. Here a basic sensitivity study on
    how basal sliding parameters in the model control the outcome discussed in
    section~4 is in order and I strongly recommend to include one in the
    manuscript. The authors can start by estimating the uncertainties in the
    chosen basal sliding parameters and run two extra simulation runs with
    their preferred climate forcing and the end member values of the
    uncertainties. This would create the most simple sensitivity study with
    respect to basal sliding, but would be extremely helpful for the argument
    made above in my general comments.}

\referee{%
    \textbf{P4156 L3-5:}
    In addition to what I have stated above on the NARR precipitation fields
    and their suitability, it is important to state at this location in the
    manuscript how the 32\,km NARR data is translated to the 10 and 5\,km
    computational grids of the current study. I disagree with the notion that a
    32\,km precipitation field can be called ``high- resolution'' in the
    context of 10 and 5\,km grid based ice sheet modelling. The input data is
    either 3 or 6 times coarser than the numerical grid, thus not at all
    high-resolution. \citet{Seguinot.etal.2014} state in their section~3.3 that
    the NARR data fields have been bilinearly interpolated to 10\,km resolution
    in their work. Did the authors do the same here for their 10 and 5\,km
    working grids? This is crucial information to be included in the
    manuscript. It has been demonstrated by spectral power analysis \citep{Jarosch.etal.2010}
    that the NARR precipitation fields do not contain any
    significant spacial information below approximately 39\,km resolution and
    that bilinear interpolation does not add any information whatsoever on
    smaller scales, which should come to no surprise. Physics-based downscaling
    techniques however are able to add spatial information to precipitation
    fields down to about 1\,km grid sizes \citep{Jarosch.etal.2010}. Taking
    these findings into the current context of the manuscript at hand, the NARR
    precipitation fields can hardly be called ``high resolution'' with their
    effective precipitation grid size of 39\,km. The authors should argue for
    their choice of not performing any downscaling whatsoever to their
    computational grids of 10 and 5\,km for precipitation and temperature and
    discuss their choice in the light of the findings from \citet{Jarosch.etal.2010}.
    Temperature however is better constrained in NARR \citep{Jarosch.etal.2010}
    and contains spectral information down to 10\,km resolution, which
    justifies the usage of NARR temperature fields on the 10\,km computational
    grid of this study. The 5\,km grid still needs to be argued for.}

@@

\referee{%
    \textbf{P4161 L10:}
    ``further analysis further;'' maybe change to ``further analysis'' or
    ``further analysis here''.}

\referee{%
    \textbf{P4166 L13:}
    double ``the'' in the sentence.}

\referee{%
    I hope the authors find my comments helpful in revising their manuscript
    and wish them success for their future endeavours.}

\begin{thebibliography}{69}

\bibitem[{Aschwanden et~al.(2012)}]{Aschwanden.etal.2012}
Aschwanden, A., Bueler, E., Khroulev, C., and Blatter, H.: An enthalpy
  formulation for glaciers and ice sheets, J. Glaciol., 58, 441--457,
  \doi{10.3189/2012JoG11J088}, 2012.

\bibitem[{Clarke et~al.(2013)}]{Clarke.etal.2013}
Clarke, G. K.~C., Anslow, F.~S., Jarosch, A.~H., Radi{\'{c}}, V., Menounos, B.,
  Bolch, T., and Berthier, E.: Ice Volume and Subglacial Topography for Western
  Canadian Glaciers from Mass Balance Fields, Thinning Rates, and a Bed Stress
  Model, J. Climate, 26, 4282--4303, \doi{10.1175/jcli-d-12-00513.1}, 2013.

\bibitem[{Jarosch et~al.(2010)Jarosch, Anslow, and Clarke}]{Jarosch.etal.2010}
Jarosch, A.~H., Anslow, F.~S., and Clarke, G. K.~C.: High-resolution
  precipitation and temperature downscaling for glacier models, Clim. Dynam.,
  38, 391--409, \doi{10.1007/s00382-010-0949-1}, 2010.

\bibitem[{Seguinot et~al.(2014)Seguinot, Khroulev, Rogozhina, Stroeven, and
  Zhang}]{Seguinot.etal.2014}
Seguinot, J., Khroulev, C., Rogozhina, I., Stroeven, A.~P., and Zhang, Q.: The
  effect of climate forcing on numerical simulations of the {C}ordilleran ice
  sheet at the {L}ast {G}lacial {M}aximum, The Cryosphere, 8, 1087--1103,
  \doi{10.5194/tc-8-1087-2014}, 2014.

\end{thebibliography}

### 2015-11-12 reviews/response-marshall.tex

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.341,0.596,0.824}
\definecolor{darkred}{rgb}{0.439,0.157,0.145}  % from Advances in Geoscience

\begin{document}
\textbf{Authors' response to S.~J.~Marshall}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\doi}[1]{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}
\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{darkred}{TODO: #1}}

@@

\referee{%
    The manuscript is well-written and beautifully illustrated, and I have very
    few substantive comments. The choices made by the authors are logical and
    well-explained, and they reach several well-substantiated conclusions: a
    two-phase Cordilleran glaciation, a reasonably robust estimate of CIS
    volume at LGM, the general model of CIS growth through multiple alpine
    icefields, and the importance of the Skeena Mountain inception centre. One
    can always quibble with specific aspects of the model design and climate
    scenarios, but the authors have explored a reasonable span of `solution
    space' and these aspects of the Cordilleran ice sheet history appear to be
    robust features of the simulations.}

\referee{%
    The modelling strategy and results presented here stand to be widely cited,
    and I expect that it will serve as a springboard for additional studies
    from others in the international community. I recommend this manuscript for
    publication in The Cryosphere without reservations.}

% ----------------------------------------------------------------------

\newpage
\sechead{1 \quad Specific comments}

\referee{%
    The Cordilleran Ice Sheet is difficult to model due to its complex
    topography and multiple inception centres (and possibly multiple
    domes/divides), strong regional climatic gradients, which require
    relatively high-resolution climate input fields, and a dearth of
    paleoclimate proxies for western North America to inform spatial and
    temporal variations in climate conditions during the glacial period. The
    authors confront these challenges well, with a adequate ice sheet model
    resolution (5 to 10\,km) and ice physics, carefully calibrated `control'
    climatology \citep[published in][]{Seguinot.etal.2014}), and a good
    exploration of different paleoclimate time series histories in this
    contribution.}

@@

\todo{Discuss in the manuscript.}

\referee{%
    Several minor points and grammatical corrections are included in the
    attached text. Nothing that will require much thought -- this is a really
    impressive piece of research, overall, and I am hard-pressed to find any
    criticism of it. It is one of the easiest reviews I have ever done.
    Congratulations to the authors and thanks for this fine work.}

\todo{Correct mistakes.}

\begin{thebibliography}{69}

\bibitem[{Seguinot et~al.(2014)Seguinot, Khroulev, Rogozhina, Stroeven, and
  Zhang}]{Seguinot.etal.2014}
Seguinot, J., Khroulev, C., Rogozhina, I., Stroeven, A.~P., and Zhang, Q.: The
  effect of climate forcing on numerical simulations of the {C}ordilleran ice
  sheet at the {L}ast {G}lacial {M}aximum, The Cryosphere, 8, 1087--1103,
  \doi{10.5194/tc-8-1087-2014}, 2014.

\end{thebibliography}

### 2015-11-12 reviews/response-stumpf.tex

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\doi}[1]{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}
\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

To A.~Stumpf,

\referee{%
    Seguinot et al. present a well designed numerical model for the Cordilleran
    Ice Sheet (CIS) in North America for the last two glaciations occurring
    over the past 120000 years. Although the constraints on such a model are
    not yet all fully understood, their proposed simulations attempt to take
    account all the complexity of the glacial system, and utilizes a variety of
    data sets. The modeling confirms what geologists have observed in the
    field; there was a fast decay of the CIS during MIS~2 and MIS~4 and
    non-glacial conditions existed during MIS (Olympia Nonglacial Interval;
    \citealp[e.g.,][]{Plouffe.Jette.1997}). I applaud them for undertaking of such a
    difficult task, and notifying the reader where input data is sparse or
    inconclusive.}

@@

\referee{%
    I provide the following general comments and observations which may help
    the authors in revising the manuscript for final publication. Many of these
    points are both my personal suggestions and also the recommendations of
    others currently researching the CIS or who have completed studies in the
    past.}

\referee{%
    1) With the large amount of research that has been undertaken to map the
    landforms and deposits of the CIS by federal and provincial scientists,
    academic faculty, and undergraduate and graduate students and determine the
    extent, volume and dynamics ice sheet, the impression left on the reader by
    the opening sentence would be incorrect. These studies have greatly
    advanced our understanding of the CIS, and could be an important dataset to
    test against the modeling. \citet{Stumpf.etal.2014} provides a list of some of
    these studies.}

@@

\referee{%
    2) To help the reader better understand the maps presented, I would
    recommend some spatial information be added (e.g., latitude/longitude
    grids; political boundaries; lakes and rivers, place names etc..).}

\referee{%
    3) Was the model tested against regional-scale ground-based data
    \citep[e.g.,][]{Ferbey.etal.2013} to constrain ice divide positions, ice flow direction,
    and ice sheet thickness?}

\referee{%
    4) For the central sector of the CIS, \citet{Stumpf.etal.2000} provides some
    insight into the chronology and effectiveness of glacial erosional during
    the MIS. In figure~5, and in the accompanying text, they describe how
    landforms on the surface formed. For example, in lake valleys east of the
    Skeena Mountains, it appeared the major glacial streamlined landforms were
    formed during a longer glacial advance phase, with ice flow paralleling the
    valleys, and later flows, some perpendicular the valley flow, only weakly
    impacted them.}

\referee{%
    5) \citet{Stumpf.etal.2000} was the first study to extensively document a
    predominant westerly directed ice flow across high elevations in the Skeena
    and Coast Mountains. This flow direction appeared to continue into the
    late-glacial period. Other subsequent studies also confirm that
    late-glacial readvance eastward out of these mountains and retreat of ice
    margins westward into these mountains was limited.}

% ----------------------------------------------------------------------

\begin{thebibliography}{69}

\bibitem[{Ferbey et~al.(2013)}]{Ferbey.etal.2013}
Ferbey, T., Arnold, H., and Hickin, A.: Ice-flow indicator compilation, British
  Columbia., British Columbia Geol. Surv., Victoria, BC, Open-File 2013-06,
  \url{http://www.em.gov.bc.ca/Mining/Geoscience/PublicationsCatalogue/OpenFiles/2013/Pages/2013-06.aspx},
  2013.

\bibitem[{Seguinot et~al.(2014)}]{Seguinot.etal.2014}
Seguinot, J., Khroulev, C., Rogozhina, I., Stroeven, A.~P., and Zhang, Q.: The
  effect of climate forcing on numerical simulations of the {C}ordilleran ice
  sheet at the {L}ast {G}lacial {M}aximum, The Cryosphere, 8, 1087--1103,
  \doi{10.5194/tc-8-1087-2014}, 2014.

\bibitem[{Stumpf et~al.(2000)}]{Stumpf.etal.2000}
Stumpf, A.~J., Broster, B.~E., and Levson, V.~M.: Multiphase flow of the late
  Wisconsinan Cordilleran ice sheet in western Canada, Geol. Soc. Am. Bull.,
  112, 1850--1863, \doi{10.1130/0016-7606(2000)112<1850:mfotlw>2.0.co;2}, 2000.

\bibitem[{Stumpf et~al.(2014)}]{Stumpf.etal.2014}
Stumpf, A.~J., Ferbey, T., Plouffe, A., Clague, J.~J., Ward, B.~C., Paulen,
  R.~C., and Bush, A.~B.: Discussion: ``Streamlined erosional residuals and
  drumlins in central British Columbia, Canada'' by J. Donald {McClenagan},
  (2013) Geomorphology 189, 41--54, Geomorphology, 209, 147--150,
  \doi{10.1016/j.geomorph.2013.10.019}, 2014.

\bibitem[{Plouffe and Jett{\'{e}}(1997)}]{Plouffe.Jette.1997}
Plouffe, A. and Jett{\'{e}}, H.: Middle Wisconsinan sediments and paleoecology
  of central British Columbia: sites at Necoslie and Nautley rivers, Can. J.
  Earth Sci., 34, 200--208, \doi{10.1139/e17-016}, 1997.

\end{thebibliography}

### 2015-11-12 reviews/response-jarosch.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage{hyperref}
\usepackage{natbib}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}
\definecolor{adgeored}{rgb}{0.439,0.157,0.145}  % from Advances in Geoscience

@@

\sechead{1 \quad General comments}

\referee{%
    Seguinot et al. present in this well written and structured manuscript a
    numerical modelling study of the Cordilleran ice sheet through the last
    glacial cycle. The model is driven by several temperature reconstructions
    based on proxy data and model output is subsequently compared in detail to
    the existing geological evidence in the region. The study is of significant
    relevance as it focuses on the Cordilleran ice sheet evolution in the past,
    which is still poorly understood.}

\referee{%
    Nevertheless, the manuscript is quite unbalanced in its presentation as it
    focuses strongly on section~4 (Comparison with geological record) and by
    doing so neglects crucial details in section~2 (Model setup). This poses a
    fundamental challenge for understanding the science presented. If it is not
    quite clear what the model does and how it performs to start with, it
    becomes difficult to discuss the results of the modelling study and why
    there are mismatches with geological evidence.}

\todo{%
    Perhaps explain here that this fundamental challenge is not one posed by
    the presentation of the manuscript, but one linked to the mutual
    understanding of the glaciological and geomorphological communities.}

@@

\referee{%
    What I advocate at this point is not a complete, strict sensitivity study
    of all parameters involved in the model setup (that would be probably a
    work package large enough to fill a science career). However several key
    parameters can be investigated with not too much effort. Contrasting the
    influence of e.g. basal sliding and ice rheology parameters with the
    influence of driving climate on the model results would help to estimate
    the overall sensitivity of the model system as well as help guiding future
    efforts performing such modelling studies. Implicitly the authors assume
    that all other model sensitivities are negligibly small in comparison to
    the driving climate. However it is obvious from an ice sheet model
    perspective that at least chosen basal sliding parameters as well as ice
    rheology parameters will strongly influence the shape and volume of the
    modelled ice sheet. Thus it would be nice to see evidence supporting the
    claim that driving climate is the only input to worry about being presented
    in the current manuscript. Or should it turn out that basal sliding and ice
    rheology play an important role too, as one would expect, then the relative
    importance of each including error estimates on the chosen parameters
    should be presented as well.}

\todo{%
    Perform new sensitivity tests on rheological parameters (water content,
    activation energies, rate factors) and sliding (friction angle). This will
    take some time. Based on the results, decide if they should be presented in
    the manuscript (before/after) discussion, in a supplement, or perhaps just
    here in this response.}

@@

\referee{%
    Generally section~4 appears to be quite long and seems to re-summarize
    known geological evidence for the region. At times the language is quite
    speculative, for example P4162 L1 and 7, P4164 L18, P4171 L7, L9, L17, L19
    and L20 and so forth. I would recommend to shorten that section to focus
    only on the geological evidence which can or can not clearly be reproduced
    by the presented model and avoid extensive speculation on what the reasons
    for mismatch are, especially in the present form of the manuscript, where a
    sensitivity study of the model itself is completely missing. However I
    leave the choice of how much geological evidence is discussed in the
    manuscript entirely up to the authors.}

\todo{%
    I have no intention to remove half the manuscript. Argue against it here.
    Geomorphology is no exact science and has to be speculative.}

### 2015-11-12 reviews/response-marshall.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage{hyperref}
\usepackage{natbib}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.341,0.596,0.824}
\definecolor{adgeored}{rgb}{0.439,0.157,0.145}  % from Advances in Geoscience

@@

\todo{%
    Logan ice core covers only the last 16\,Ka and is not a good proxy for
    temperature. The Vancouver SST record covers only deglaciation. Check new
    papers for updates. Discuss in the manuscript.}

\referee{%
    Similarly it is difficult to know the errors and uncertainties associated
    with the assumption of fixed modern-day spatial patterns for temperature
    and precipitation. I suspect that the sensitivity of this assumption far
    exceeds that associated with the different paleoclimate proxies. Such that,
    for example, one could readily imagine different assumptions, such as a
    maritime effect that gives reduced glacial cooling near the coast vs. in
    the interior, that is a stronger effect than the difference between
    different paleo-climate proxies with respect to the timing of LGM, ice
    divide structure, etc. But this is a very reasonable start, what the
    authors have done -- there is always going to be more parameter space to
    explore in future studies. As above, I would perhaps just suggest a small
    discussion of the authors' opinion on this question, the uncertainty or
    possible influence of this assumption of modern-day climate patterns.}

\todo{%
    Perhaps we forgot to mention that as a limitation in the manuscript. Find
    the appropriate place to add this in the discussion, and rephrase the above
    points into the manuscript.}
